Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-01-15 paperIEEE.tex
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@82 db7c848b-7755-4df2-b763-8796797afa52

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
  \begin{abstract}
    Developing complex, real world graphics applications which utilize multiple
    GPUs and computers for interactive 3D rendering tasks is a complex task, it
    requires expertise in distributed systems and parallel rendering in addition
    to the application domain itself. We present a mature parallel rendering
    framework which provides a large set of features, algorithms and system
    integration for a wide range of real-world research and industry
    applications. Based on the basic architecture and implementation of the
    \textsf{Equalizer} parallel rendering framework presented in \cite{EMP:09},
    we show how a variety of generic algorithms are integrated in the framework
    to help application scalability and development in many different domains,
    highlight how concrete applications benefit from the diverse aspects and use
    cases of \textsf{Equalizer}, and show new experimental results for parallel
    rendering and data distribution.
  \end{abstract}

@@

%----------------------------------------------------------------------
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.
\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
%----------------------------------------------------------------------

@@

%------------------------------------------------------------------------------
\section{Performance}
%------------------------------------------------------------------------------

@@

The initial version of \textsf{Equalizer} implemented the basic sort-first (2D),
sort-last (DB), stereo (EYE) and multilevel decompositions \cite{EMP:09}. In the
following we present the newly added decomposition modes and motivate their use
case, which bring the overall feature set way beyond the typical sort-first and
sort-last rendering modes. \fig{fig:compounds} provides an overview of the new
modes. We describe how to setup compound descriptions in \cite{EMP:09}.

@@

Tile (\fig{fig:tiles}) and chunk decompositions are a variant of sort-first and
sort-last rendering, respectively. They decompose the scene into a predefined
set of fixed-size image tiles or database ranges. These tasks are queued and
processed by all source channels by polling a server-central queue. Prefetching
ensures that the task communication overlaps with rendering. As shown in~
\cite{SPEP:16}, these modes provide better performance due to being inherently
load-balanced, as long as there is an insignificant overhead for the render task
setup. This mode is transparent to \textsf{Equalizer} applications.

@@

%------------------------------------------------------------------------------
\section{Virtual Reality}
%------------------------------------------------------------------------------

@@

\subsection{Head Tracking}

@@

\subsection{Dynamic Focus Distance}

@@

\subsection{Asymmetric Eye Position}

@@

\subsection{Model Unit}

@@

\subsection{Runtime Stereo Switch}

@@

%------------------------------------------------------------------------------
\section{Usability}
%------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------
\section{Applications\label{sApplications}}
%----------------------------------------------------------------------

@@

\subsection{Livre}

@@

\subsection{RTT Deltagen}

@@

\subsection{RTNeuron}

@@

\subsection{Raster}

@@

\subsection{Bino}

@@

\subsection{Omegalib}

@@

\subsection{Decomposition Modes}

@@

\subsection{Object Distribution}

@@

%----------------------------------------------------------------------
\appendices
% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi
%----------------------------------------------------------------------
We would like to thank and acknowledge the following institutions and projects
for providing the 3D geometry and volume test data sets: the Digital
Michelangelo Project, Stanford 3D Scanning Repository, Cyberware Inc.,
volvis.org and the Visual Human Project.  This work was partially supported by
the Swiss National Science Foundation Grant 200021-116329/1.\FIXME{SNF and
Hasler as well as EU grants}

@@

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

@@

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

>>>>

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
  \begin{abstract}
    Developing complex, real world graphics applications which leverage multiple
    GPUs and computers for interactive 3D rendering tasks is a complex task. It
    requires expertise in distributed systems and parallel rendering in addition
    to the application domain itself. We present a mature parallel rendering
    framework which provides a large set of features, algorithms and system
    integration for a wide range of real-world research and industry
    applications. Based on the basic architecture and implementation of the
    \textsf{Equalizer} parallel rendering framework presented in \cite{EMP:09},
    we show how a variety of generic algorithms can be integrated in the framework
    to help application scalability and development in many different domains,
    highlighting how concrete applications benefit from the diverse aspects and use
    cases of \textsf{Equalizer}, and we present new experimental results for parallel
    rendering and data distribution.
  \end{abstract}

@@

%----------------------------------------------------------------------
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.
\IEEEraisesectionheading{\section{Introduction} \label{sec:introduction}}
%----------------------------------------------------------------------

@@

%------------------------------------------------------------------------------
\section{Performance Features}
%------------------------------------------------------------------------------

@@

The initial version of \textsf{Equalizer} implemented the basic sort-first (2D),
sort-last (DB), stereo (EYE) and multilevel decompositions \cite{EMP:09}. In the
following we present the newly added decomposition modes and motivate their use
case, which bring the overall feature set way beyond the typical sort-first and
sort-last rendering modes. \fig{fig:compounds} provides an overview of the new
modes. The way how compounds are setup is still as given in \cite{EMP:09}.

@@

Tile (\fig{fig:tiles}) and chunk decompositions are a variant of sort-first and
sort-last rendering, respectively. They decompose the scene into a predefined
set of fixed-size image tiles or database ranges. These tasks, or work packages are queued and
processed by all source channels by polling a server-central queue. Prefetching
ensures that the task communication overlaps with rendering. As shown in~
\cite{SPEP:16}, these modes can provide better performance due to being inherently
load-balanced, as long as there is an insignificant overhead for the render task
setup. This mode is transparent to \textsf{Equalizer} applications.

@@

%------------------------------------------------------------------------------
\section{Virtual Reality Features}
%------------------------------------------------------------------------------

@@

%------------------------------------------------------------------------------
\subsection{Head Tracking}

@@

%------------------------------------------------------------------------------
\subsection{Dynamic Focus Distance}

@@

%------------------------------------------------------------------------------
\subsection{Asymmetric Eye Position}

@@

%------------------------------------------------------------------------------
\subsection{Model Unit}

@@

%------------------------------------------------------------------------------
\subsection{Runtime Stereo Switch}

@@

%------------------------------------------------------------------------------
\section{Usability Features}
%------------------------------------------------------------------------------

@@

%----------------------------------------------------------------------
\section{Example Applications} \label{sApplications}
%----------------------------------------------------------------------

@@

%------------------------------------------------------------------------------
\subsection{Livre}

@@

%------------------------------------------------------------------------------
\subsection{RTT Deltagen}

@@

%------------------------------------------------------------------------------
\subsection{RTNeuron}

@@

%------------------------------------------------------------------------------
\subsection{RASTeR}

@@

%------------------------------------------------------------------------------
\subsection{Bino}

@@

%------------------------------------------------------------------------------
\subsection{Omegalib}

@@

%------------------------------------------------------------------------------
\subsection{Decomposition Modes}

@@

%------------------------------------------------------------------------------
\subsection{Object Distribution}

@@

%----------------------------------------------------------------------
\appendices
% use section* for acknowledgment
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi
%----------------------------------------------------------------------
We would like to thank and acknowledge the following institutions and projects
for providing the 3D geometry and volume test data sets: the Digital
Michelangelo Project, Stanford 3D Scanning Repository, Cyberware Inc.,
volvis.org and the Visual Human Project. 
%
This work was partially supported by
the Swiss National Science Foundation Grants 200021-116329 and 200020-129525,
the Swiss Commission for Technology and Innovation CTI/KTI Project 9394.2 PFES-ES,
the EU FP7 People Programme (Marie Curie Actions) under REA Grant Agreement n$^{\circ}290227$ and
a Hasler Stiftung grant (project number $12097$).

@@

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{David Steiner}
Biography text here.
\end{IEEEbiographynophoto}

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/renato}}]{Prof. Dr. Renato Pajarola}
has been a Professor in computer science at the University of Zurich since 2005, leading the Visualization and MultiMedia Lab (VMML). He has previously been an Assistant Professor at the University of California Irvine and a Postdoc at Georgia Tech. He has received his Dipl. Inf-Ing. ETH and Dr. sc. techn. degrees in computer science from the Swiss Federal Institute of Technology (ETH) Zurich in 1994 and 1998 respectively. He is a Fellow of the Eurographics Association and a Senior Member of both ACM and IEEE. His research interests include real-time 3D graphics, interactive data visualization and geometry processing.
\end{IEEEbiography}
